Tidal field anisotropy as a tracer of cosmic voids

### 2015-02-11 cosmic_voids.tex
intro revisada

%=========================================================================
%		FRONT MATTER
%=========================================================================
\title{Tidal field anisotropy as a tracer of cosmic voids}
\author[S. Bustamante and J.E. Forero-Romero]{
\parbox[t]{\textwidth}{\raggedright 
  Sebastian Bustamante \thanks{sebastian.bustamante@udea.edu.co}$^{1}$,
  Jaime E. Forero-Romero$^{2}$ 
}
\vspace*{6pt}\\
$^1$Instituto de F\'{\i}sica - FCEN, Universidad de Antioquia, Calle
67 No. 53-108, Medell\'{\i}n, Colombia\\ 
$^2$Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia
}

@@

\begin{abstract}
Finding and characterizing underdense regions (voids) in the large
scale structure of the Universe is an important task in cosmological
studies.  
In this paper we present a novel approach to find voids in
cosmological simulations.  
Our approach is based on algorithms that use the tidal and the
velocity shear tensors to locally define the cosmic web.
Voids are identified using the fractional anisotropy (FA) computed
from the eigenvalues of each web scheme. 
We define the void boundaries using a watershed transform based on the
local minima of the FA and its boundaries as the regions where the FA
is maximized
This void identification technique does not have any free parameters
and does not make any assumption on the shape or structure of the
voids.  
We test the method on the Bolshoi simulation and report on the density
and velo ity profiles for the voids found using this new scheme. 
\end{abstract}

@@

Since voids were found in the first compiled galaxy surveys 
they have been identified as one of the most striking features of the
Cosmic Web \citep{Chincarini75,  Gregory78, Einasto80M, Einasto80N,
  Kirshner81, Zeldovich82,Kirshner87, Bond96}. 
However,  due to the large volume extension of void regions ($\sim
5-10\ \mbox{Mpc}  h^{-1}$), statistically meaningful catalogues of
voids \citep{Pan10,  Sutter12b, Nadathur14} have only become available
through  modern galaxy surveys like the two-degree field Galaxy
Redshift Survey \citep{ Colless01, Colless03} and the Sloan Digital
Sky Survey \citep{York00, Abazajian03}.
This advancement generated a great interest to study voids
observationally during the last decade \citep{Hoyle04, Croton04, Rojas05,
  Ceccarelli06, Patiri06a, Tikhonov06, Patiri06b,Tikhonov07,
  BendaBeckmann08, Foster09, Ceccarelli13, Sutter14a}.

On the theoretical side the rough theoretical framework that explains
the origin of voids was established in the seminal work of
\citet{Zeldovich70} and refined in the following decades.  
The first detailed theoretical models describing formation, dynamics
and properties of  voids \citep{Hoffman82, Icke84, Bertschinger85,
  Blumenthal92} were  complemented and extended by numerical studies
\citep{Martel90, Regos91, Weygaert93, Dubinski93, Bond96}. 
Currently, the dominant tendency to study voids relies on using data
from N-body simulations. For an extensive compilation of previous 
numerical works we refer the reader to \citet{Colberg08}.

The relevance of voids to cosmological studies can be summarized in
three aspects \citep{Platen07}. 
First, voids are a key ingredient of the Cosmic Web. 
They dominate the volume distribution at large-scales and
additionally, they compensate overdense structures in the total budget
of  matter.  
Second, voids provide a valuable resource for inferring and probing 
cosmological parameters as their structure and dynamics are highly 
determined by these values. 
Third, they are a  largely a largely pristine environment to test
galaxy evolution.

Although a visual recognition of voids in galaxy surveys and simulations
is possible in most cases, a formal systematic identification is desirable
for statistical studies. 
However, the community does not have an unambiguous definition of cosmic
voids. 
There are nany different void finding techniques in the literature
(for a detailed comparison of different schemes,  see the Void Finder
Comparison Project \citet{Colberg08}). 
In spite of the diversity of existing schemes, they can be roughly
classified into two main types. First, geometric schemes based on
point spatial or redshift  distribution of galaxies in surveys or dark
matter halos in  simulations \citep{Kauffmann91, Muller00,
  Gottlober03, Hoyle04, Brunino07,  Foster09, Micheletti14, Sutter14};
second, schemes based on the smooth and continuous matter density
field either from simulations or from reconstruction procedures on
surveys \citep{Plionis02, Colberg05,  Shandarin06, Platen07,
  Neyrinck08, MunozCuartas11, Neyrinck13, Ricciardelli2013}.

Here we introduce a new tracer of the structure of cosmic voids that
can be used once the continous matter density or velocity distribution
is known. 
The algorithm is based on two tensorial schemes used to classify the
cosmic web. 
The first (the T-web) is based on the Hessian of the gravitational potential or
tidal tensor \citep{Hahn07, Forero09}. 
The second (the V-web) is based on the velocity shear tensor
\citep{Hoffman12}. 
These web schemes classify the Cosmic Web into four different types of
environment depending on the counting of the number of eigenvalues below an 
user-defined threshold ($\lambda_{rm th}$): voids, sheets, filaments and
knots for $3,2,1$ or $0$ eigenvalues below $\lambda_{\rm th}$, respectively.

This procedure allows a new description of the internal structure
of voids that goes beyond a simple definition of a void as just an
underdense regions in the large-scale matter distribution.
The tidal and the shear tensors encode more information than the
density field as they trace the collapsing or expanding
nature of the matter field, which defines the dynamics of the Cosmic
Web.

The tracer that we use to define the voids is the fractional
anisotropy (FA) computed from the set of eigenvalues of the tensor
under consideration. 
The FA was initially introduced by \citet{Basser95} to quantify the
anisotropy degree of the diffusivity of water molecules through
cerebral tissue in nuclear magnetic resonance
imaging.
\citet{Libeskind13} introduced this concept in the context of Cosmic
Web classification schemes.

Once we establish the FA as a void tracer, we proceed to
identify individual voids as basins of local minima. 
For this we implement a \textit{watershed transform algorithm}
\citep{Beucher79,Beucher93} which has been used to define as catching
basins of local minima of the density field
\citep{Platen07,Neyrinck08}; we use insted minima of the FA field to
find define the voids.

This paper is organized as follows. 
In Section \ref{sec:algorithms_cosmic_web} we describe the algorithms
we use to find the cosmic web in N-body simulations. In Section 
\ref{sec:simulations} we describe the used simulation.
Then in Section \ref{sec:bulk_voids} we explain in detail our void
finder based on the fractional anisotropy of the tidal and shear
tensor fields.
%In Section \ref{sec:experiments} we describe the numerical experiments
%we perform to show the performance of the new void finding technique.
Our results are presented in \ref{sec:results} to finally comment and
conclude about these findings in Section \ref{sec:conclusions}.

>>>>

%=========================================================================
%		FRONT MATTER
%=========================================================================
\title{Tidal field anisotropy as a tracer of cosmic voids}
\author[S. Bustamante and J.E. Forero-Romero]{
\parbox[t]{\textwidth}{\raggedright 
  Sebastian Bustamante \thanks{sebastian.bustamante@udea.edu.co}$^{1}$,
  Jaime E. Forero-Romero \thanks{je.forero@uniandes.edu.co}$^{2}$ 
}
\vspace*{6pt}\\
$^1$Instituto de F\'{\i}sica - FCEN, Universidad de Antioquia, Calle
67 No. 53-108, Medell\'{\i}n, Colombia\\ 
$^2$Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia
}

@@

\begin{abstract}
Finding and characterizing underdense regions (voids) in the large
scale structure of the Universe is an important task in cosmological
studies.  
In this paper we present a novel approach to find voids in
cosmological simulations.  
Our approach is based on algorithms that use the tidal and the
velocity shear tensors to locally define the cosmic web.
Voids are identified using the fractional anisotropy (FA) computed
from the eigenvalues of each web scheme. 
We define the void boundaries using a watershed transform based on the
local minima of the FA and its boundaries as the regions where the FA
is maximized.
This void identification technique does not have any free parameters
and does not make any assumption on the shape or structure of the
voids.  
We test the method on the Bolshoi simulation and report on the density
and velocity profiles for the voids found using this new scheme. 
We find that...
\end{abstract}

@@

Voids have been identified as one of the most striking features of the
Universe on its larges scales since they were found in the first
galaxy surveys \citep{Chincarini75, Gregory78, Einasto80M, Einasto80N,
  Kirshner81, Zeldovich82,Kirshner87}.  
However, due to the large volume extension of void regions ($\sim
5-10\ \mbox{Mpc}  h^{-1}$), statistically meaningful catalogues of
voids \citep{Pan10,  Sutter12b, Nadathur14} have only become available
through modern galaxy surveys such as the two-degree field Galaxy
Redshift Survey (2dF) \citep{ Colless01, Colless03} and the Sloan
Digital Sky Survey (SDSS)\citep{York00, Abazajian03}.
These observational breakthughs generated a great interest in the last
decade to study voids \citep{Hoyle04, Croton04, Rojas05,
  Ceccarelli06, Patiri06a, Tikhonov06, Patiri06b,Tikhonov07,
  BendaBeckmann08, Foster09, Ceccarelli13, Sutter14a}.

On the theoretical side the basic framework that explains
the origin of voids was established in the seminal work of
\citet{Zeldovich70} and refined in the following decades.  
The first detailed theoretical models describing formation, dynamics
and properties of  voids \citep{Hoffman82, Icke84, Bertschinger85,
  Blumenthal92} were  complemented and extended by numerical studies
\citep{Martel90, Regos91, Weygaert93, Dubinski93, Bond96}. 
Currently, the most popular approach to study voids relies on N-body
simulations. For an extensive compilation of previous  
numerical works we refer the reader to \citet{Colberg08}.

The relevance of voids to cosmological studies can be summarized in
three aspects \citep{Platen07}. 
Firstly, voids are a key ingredient of the Cosmic Web. 
They dominate the volume distribution at large scales and
additionally, compensating overdense structures in the total matter
budget. 
Secondly, voids provide a valuable resource to estimate  
cosmological parameters as their structure and dynamics are sensitive
to them. 
Finally, they are a largely pristine environment to test galaxy
evolution.

Although visual recognition of voids in galaxy surveys and simulations
is possible in most cases, we need a clear algorithmic identification
procedure to make statistical studies.
Nevertheless, the community has not reached yet an unambiguous
definition of cosmic voids, with many different void finding
techniques in the literature  (for a detailed comparison of different
schemes,  see the publication on the results of the Void Finder
Comparison Project \citet{Colberg08}).  
In spite of the diversity of existing schemes, they can be roughly
classified into two types. First, geometric schemes based on
point distributions (either real or redshift space) of galaxies in
surveys or dark matter halos in  simulations \citep{Kauffmann91, Muller00, 
  Gottlober03, Hoyle04, Brunino07,  Foster09, Micheletti14, Sutter14}.
Second, schemes based on the smooth and continuous matter density
field either from simulations or from reconstruction procedures on
surveys \citep{Plionis02, Colberg05,  Shandarin06, Platen07,
  Neyrinck08, MunozCuartas11, Neyrinck13, Ricciardelli2013}. 
Our work is based on the second kind of schemes.

Here we introduce a new algorithm to define voids over the continous
matter density or velocity distribution defined on a fixed and
homogeneous spatial grid.
The algorithm is based on two tensorial schemes used to classify the
cosmic web. 
The first (the T-web) is based on the Hessian of the gravitational potential or
tidal tensor \citep{Hahn07, Forero09}. 
The second (the V-web) is based on the velocity shear tensor
\citep{Hoffman12}. 
Our procedure allows a description of the internal structure
of voids that goes beyond a simple definition of a void as just an
underdense regions in the large-scale matter distribution.
The tidal and the shear tensors encode more information than the
density field as they trace the collapsing or expanding
nature of the matter field, which defines the dynamics of the Cosmic
Web.

The tracer that we use to define the voids is the fractional
anisotropy (FA) computed from the set of eigenvalues of the tensor
under consideration. 
The FA was initially introduced by \citet{Basser95} to quantify the
anisotropy degree of the diffusivity of water molecules through
cerebral tissue in nuclear magnetic resonance
imaging and \citet{Libeskind13} introduced this concept in the context
of Cosmic Web classification schemes.

Once we establish the FA as a void tracer, we proceed to
identify individual voids as basins of FA local minima. 
At this point we implement a \textit{watershed transform algorithm}
\citep{Beucher79,Beucher93} which has been used to define voids as
catching basins of local minima of the density field
\citep{Platen07,Neyrinck08}.

This paper is organized as follows. 
In Section \ref{sec:algorithms_cosmic_web} we describe the algorithms
we use to find the cosmic web in N-body simulations. 
In Section  \ref{sec:simulations} we describe the N-body simulation we
use to test our algorithm. 
In Section \ref{sec:bulk_voids} we explain in detail our void finder
based on the fractional anisotropy of the tidal and shear tensor
fields. 
Our results are presented in \ref{sec:results} to finally comment and
conclude about these findings in Section \ref{sec:conclusions}.
